---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# survtrans

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/survtrans)](https://CRAN.R-project.org/package=survtrans)
[![R-CMD-check](https://github.com/ziyangg98/survtrans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ziyangg98/survtrans/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ziyangg98/survtrans/graph/badge.svg)](https://app.codecov.io/gh/ziyangg98/survtrans)
<!-- badges: end -->

**survtrans** provides transfer learning for survival analysis using Cox proportional hazards models. It transfers survival information from source domain(s) to a target domain with three-layer penalization:

- **Sparse** (lambda1): variable selection on target coefficients
- **Local** (lambda2): shrinks source-target coefficient differences, encouraging shared effects
- **Prior transfer** (lambda3): incorporates prior knowledge about which source groups are informative via a customizable weight matrix

## Installation

You can install the development version of **survtrans** with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("ziyangg98/survtrans")
```

## Example

Fit a transfer learning Cox model on simulated data with 5 groups (1 target + 4 sources, 20 features, true support on X1--X4):

```{r example}
library(survtrans)

formula <- Surv(time, status) ~ . - group - id
fit <- coxtrans(
  formula, sim2, sim2$group, 1,
  lambda1 = 0.075, lambda2 = 0.04, lambda3 = 0.04, penalty = "SCAD"
)
summary(fit)
```

The model correctly identifies the three-layer structure: X1 and X2 are transferred via the prior constraint, X3 and X4 are shared across all groups via local shrinkage, and X5--X20 are sparse (zero).

### Custom prior matrix

When source groups have known structure (e.g., tissue type), you can define a prior matrix to encode which sources are informative for the target:

```{r prior}
pm <- rbind(
  tissue_A = c(0.5, 0.5, 0, 0),
  tissue_B = c(0, 0, 0.5, 0.5)
)
colnames(pm) <- c("2", "3", "4", "5")

fit2 <- coxtrans(
  formula, sim2, sim2$group, 1,
  lambda1 = 0.075, lambda2 = 0.04, lambda3 = c(0.04, 0.04),
  prior_matrix = pm, penalty = "SCAD"
)
summary(fit2)
```

### Automatic tuning

`cv.coxtrans()` selects all three penalty parameters jointly via K-fold
cross-validation over a full `lambda1 × lambda2 × lambda3` grid, minimising
the held-out partial likelihood deviance. It supports both the **lambda.min**
rule (minimum CV deviance) and the **lambda.1se** rule (most sparse model
within one standard error of the minimum, consistent with **glmnet**). It
returns a `cv.coxtrans` object with CV diagnostics and the final refitted
models at both rules.

```{r cv}
cv_fit <- cv.coxtrans(
  formula, sim2, sim2$group, target = 1, penalty = "SCAD", ncores = 8
)
cv_fit
```

The CV curve along the `lambda1` axis (with `lambda2` and `lambda3` fixed at
their optimal values) can be visualised with `plot()`:

```{r cv-plot, fig.alt = "Cross-validation curve for cv.coxtrans"}
plot(cv_fit)
```

Access the final fitted model via `$coxtrans.fit` (lambda.min) or
`$coxtrans.fit.1se` (lambda.1se). Use `coef(cv_fit, s = "lambda.1se")` or
`predict(cv_fit, s = "lambda.1se", ...)` to extract from the sparser model.

```{r cv-summary}
summary(cv_fit$coxtrans.fit)
```

### Baseline hazard

```{r, basehaz, fig.alt = "Estimated vs. True Cumulative Baseline Hazard Function"}
library(ggplot2)

basehaz_pred <- basehaz(fit)
basehaz_pred$color <- ifelse(
  as.numeric(basehaz_pred$strata) %% 2 == 0, "Group 2", "Group 1"
)
ggplot(
  basehaz_pred,
  aes(
    x = time,
    y = basehaz,
    group = strata,
    color = factor(color),
    linetype = "Estimates"
  )
) +
  geom_line() +
  geom_line(
    aes(x = time, y = time^2 / 2, color = "Group 1", linetype = "True")
  ) +
  geom_line(
    aes(x = time, y = time^3 / 3, color = "Group 2", linetype = "True")
  ) +
  labs(
    title = "Cumulative Baseline Hazard Function (Estimated vs. True)",
    x = expression(t),
    y = expression(Lambda[0](t))
  ) +
  scale_linetype_manual(values = c("Estimates" = "dashed", "True" = "solid")) +
  guides(
    color = guide_legend(title = "Strata"),
    linetype = guide_legend(title = "Type")
  )
```
